Functory: Une bibliothèque de calcul distribué \\ pour Objective Caml\footnotetext{Ce travail a été réalisé en partie dans le cadre du projet U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2010-10-21 doc/jfla-2011/main.tex
*** empty log message ***

\paragraph{N-reines.}\label{sec:n-queens}

The next example is the classical $N$-queens problem, where we
compute the total number of ways to place $N$ queens on a $N\times N$
chessboard in such a way no two queens attack each other.
We use a standard backtracking algorithm for this problem, which
places the queens one by one starting from the first row.
Distributing the computation is thus quite easy: we consider all
possible ways to place queens on the first $D$ rows and then perform
the subsequent search in parallel. Choosing $D=1$ will result in
exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
so on.
% TODO: make a picture to illustrate positioning of 2 queens on the
% first 2 rows

Each task only consists of three integers and its result is one integer,
which is the total number of solutions for this task.
We make use of function \of{map_local_fold} from the derived API,
where \of{map} is performing the search and \of{fold} simply adds the
intermediate results.
In the network configuration, we make use of the \of{Network.Same}
module, workers and master being the same executable.

Drawing the Mandelbrot set is another classical example that could be
distributed easily, since the color of each point can be computed
independently of the others. 
Let us assume we are given the coordinates of the region to be drawn,
along with the width \of{w} and height \of{h} of the resulting image
(in pixels). 
Let us assume the total number of tasks $\of{t} \ge 1$ is given as a
parameter. It is immediate to split the image into \of{t} sub-images,
each of which is computed in parallel with and independently of the
others. For instance, the image could be split into horizontal slices
or, more generally, into rectangular blocks.

Each task is thus four floating-point numbers denoting the region
coordinates, together with two integers denoting the dimensions of the
sub-image to be drawn. The result of the task is a matrix of pixels,
of size $(\of{w}\times\of{h}) / \of{t}$. 
For instance, drawing a $800\times 600$ image using 20 tasks will
result in 20 sub-images of size $176,000$ bytes each,
assuming each pixel is encoded in four bytes.

In the network configuration, we deliberately choose to have two
different programs for master and workers, using the \of{Network.Poly}
module (actually, a single source code with a command line option).

% \subsection{Polynomial Multiplication}
% TODO?

\subsection{SMT Solvers}\label{sec:SMT}

Here we demonstrate the potential of our library for our application
needs as mentioned in the introduction. In this case study,
we consider 80 challenging verification conditions
(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
VC is stored in a file, which is accessible over
NFS. The purpose of the experiment is to check the validity of each VC
using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
which are all installed on each of the machines in use.

The master program proceeds by reading the file names, turning them
into tasks by multiplying them by the number of provers (more than
300 tasks in total).
Each worker in turn invokes the given prover on the given file, within
a timeout limit.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{unknown} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.

The result of each computation is a pair denoting the status and the
time spent in the prover call. The master collects these results and
sums up the timings for each prover and each possible status.

>>>>

\paragraph{N-reines.}\label{sec:n-queens}
Il s'agit du problème classique consistant à calculer le nombre de
façons de disposer $N$ reines sur un échiquier $N\times N$ sans que
deux d'entre elles soient en prise.
On utilise un algorithme standard, consistant à positionner une reine
sur chaque ligne de l'échiquier, en partant de la première ligne. Il
est facile de distribuer le calcul: on considère toutes les façons de
placer les reines des $D$ premières lignes et on effectue le reste du
calcul en parallèle. Pour $D=1$ on obtient ainsi $N$ tâches ; pour
$D=2$ on obtient $N^2-3N+2$ tâches ; et ainsi de suite.
Chaque tâche est composée de quelques entiers et son résultat est un
entier donnant le nombre de solutions pour cette tâche.
On utilise la fonction \of{map_local_fold} où \of{map} effectue la
recherche et \of{fold} somme les résultats intermédiaires.

\paragraph{L'ensemble de Mandelbrot.}
Dessiner l'ensemble de Mandelbrot est un autre exemple de calcul
aisément distribuable, puisque la couleur de chaque point peut être
calculée indépendamment. Supposons donnée une région à dessiner, ainsi
que la taille $\of{w} \times \of{h}$ en pixels de l'image finale.
On se donne un nombre de tâches $\of{t} \ge 1$ comme paramètre. Il est
immédiat de découper l'image en \of{t} sous-images, par exemple en
bandes horizontales (si $\of{h} \ge \of{t}$) ou plus généralement en
blocs rectangulaires.
Chaque tâche est constituée de quatre nombre flottants définissant la
région à dessiner, ainsi que deux entiers donnant la taille de l'image
correspondante. Le résultat est une matrice de pixels, de taille
$(\of{w}\times\of{h}) / \of{t}$.
Ainsi, dessiner une image $800\times 600$ en utilisant 20 tâches
donnera 20 sous-images de $176\,000$ octets chacune, en supposant
chaque pixel représenté par quatre octets.

\paragraph{Démonstrateurs automatiques.}\label{sec:SMT}
Le dernier exemple correspond à celui mentionné dans l'introduction.
Il s'agit ici de vérifier la validité de 80 obligations de preuve
(OP), issues de la plate-forme Why~\cite{filliatre07cav}, à l'aide de
quatre démonstrateurs automatique de la famille SMT (à savoir
Alt-Ergo, Simplify, Z3 et CVC3).  Chaque OP est vérifiée avec chaque
démonstrateur, ce qui fait un total de 320 tâches.
Les OP sont contenues dans des fichiers accessibles par NFS et une
tâche est donc un nom de fichier et un nom de démonstrateur. 
Le démonstrateur est appelé comme un programme externe, avec un temps
d'exécution maximal. Le
résultat d'une tâche est la réponse du démonstrateur (valide, abandon,
temps limite atteint, erreur pendant l'exécution) et son temps de calcul.
Le patron collecte les résultats et les présente au final sous forme
d'une table synthétique.
